---
title: "Bushfire Risk in Southeastern Australia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bushfire-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## [Introduction]{.underline}

The 2019–20 "Black Summer" bushfires were unprecedented in scale and severity.\
This vignette reproduces the analysis of Oldenborgh et al. (2021, *NHESS*), showing how climate change has increased the likelihood of extreme fire weather in southeastern Australia.\
It demonstrates the `bushfire` dataset, functions, and interactive app included in this package.

## [The dataset]{.underline}

The dataset `bushfire` contains yearly summaries of the Fire Weather Index (FWI) for southeastern Australia, September–February fire seasons, 1979–2020.

```{r}
library(bushfire)
head(bushfire)
```

## [Plotting fire weather trends]{.underline}

We can visualise long‑term changes in fire weather indices using the function plot_fwi_trend().

```{r}
plot_fwi_trend(bushfire, metric = "mean")
plot_fwi_trend(bushfire, metric = "max")
```

These plots show upward trends in both mean and maximum FWI, consistent with the observed increase in extreme fire weather.

## [Interactive exploration]{.underline}

The package also includes a Shiny app for interactive analysis. Users can filter years, switch between mean and max FWI, and view both plots and summary tables. shiny::runApp(system.file("app", package = "bushfire"))

## [Conclusion]{.underline}

This vignette confirms that extreme fire weather has become more likely in recent decades, consistent with attribution studies. The package provides a reproducible workflow: dataset, documented functions, and an interactive app for exploring bushfire risk.

## [References]{.underline}

-   van Oldenborgh, G. J., et al. (2021). Attribution of the Australian bushfire risk to anthropogenic climate change. NHESS, 21(3), 941–960.
